---
title: 'Easy Way #2'
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This example uses the ipeds library to carry out a simple regression analysis involving data from about half a dozen different IPEDS surveys from the same year. It is intended to investigate the impact of a quarter calendar system on graduation rates. It is a rather silly analysis, however, and should be regarded as  demonstration of what *can* be done with the package, not perhaps what *should* be done with the package. 

We’ll use these IVs in our analysis:

* Size of graduation cohort
* Selectivity of the institution
* Tuition $
* Control (public/private)
* Locale (city/town/suburb/rural)
* Student:faculty ratio
* Calendar system (semester/quarter)


```{r}
library(ipeds)
library(tidyverse)
library(gvlma)
library(car)
library(reshape2)

dir <- "C:\\Users\\mjs26\\Documents\\data\\downloaded"
min_school_size <- 100
```

# What's available?

Helpful for a quick reminder of the various IPEDS surveys and their abbreviations. 

```{r}

data(surveys)
surveys %>% select(c('SurveyID','Survey','Title'))

```

# Get the data

We're going to grab the survey files one at a time, merging (joining) them together by unit id as we go. The three IC files are first up:

```{r}
directory <- ipeds_survey(table='HD',year=2017, dir=dir)
names(directory) <- tolower(names(directory))

charges <- ipeds_survey('IC_AY', year=2017, dir=dir)
names(charges) <- tolower(names(charges))

charges = charges[,c('unitid',
 'tuition1', 'fee1', 'hrchg1',  #In-district average tuition for full-time undergraduates
 'tuition2', 'fee2', 'hrchg2',  #In-state average tuition for full-time undergraduates
 'tuition3', 'fee3', 'hrchg3',  #Out-of-state average tuition for full-time undergraduates
 'tuition5', 'fee5', 'hrchg5', #In-district average tuition full-time graduates
 'tuition6', 'fee6', 'hrchg6', #In-state average tuition full-time graduates
 'tuition7', 'fee7', 'hrchg7')] #Out-of-state average tuition full-time graduates

dirCharges = merge(charges, directory, by='unitid', all.x=TRUE)

ic <- ipeds_survey(table='IC',year=2017, dir=dir)
names(ic) <- tolower(names(ic))

dirCharges <- merge(dirCharges, ic, by='unitid', all.x=TRUE)


```

Then Admissions:

```{r}

admissions <- ipeds_survey(table='ADM',year=2017, dir=dir)
names(admissions) <- tolower(names(admissions))

```

Graduation rates:

```{r}

gradrates <- (ipeds_survey('GR',year=2017, dir=dir))
names(gradrates) <- tolower(names(gradrates))
gradrates <- gradrates[which(gradrates$grtype %in% c(2,3)),]

# extract the 150% graduation rate
theRates <- dcast(gradrates, unitid ~ grtype, value.var = 'grtotlt')
names(theRates) <- c('unitid','adjusted_cohort','completers')
theRates$rate <- theRates$completers/theRates$adjusted_cohort
```

Eliminate any schools with missing graduation rates:

```{r}

theRates <- theRates[which(!is.na(theRates$rate)),]

```



And any with less than `r min_school_size` in the grad rate cohort


```{r}

d1 <- merge(dirCharges, theRates, by='unitid', all.y=TRUE)
d1 <- d1[which(d1$calsys %in% c(1,2)),]
d1$calsys <- as.factor(d1$calsys)
levels(d1$calsys) <- c('Semester','Quarter')
d1 <- d1[which(d1$adjusted_cohort > min_school_size),]

```

IPEDS Admissions gives us selectivity.

```{r}
d1 <- merge(d1, admissions, by='unitid', all.x=TRUE)
d1$select <- d1$admssn / d1$applcn
```

Here's Fall Enrollment, which is where student:faculty ratio lives.

```{r}

fallenr <- ipeds_survey(table='EFD', year=2017, dir=dir)
names(fallenr) <- tolower(names(fallenr))
d1 <- merge(d1, fallenr, by='unitid', all.x=TRUE)
d1 <- d1[which(!is.na(d1$stufacr)),] # remove any schools with missing s:f ratio
```


That's all the data we need. Do our continuous variables have sensible shapes?

```{r}

plot(density(d1$adjusted_cohort), main="Cohort")
plot(density(d1[which(!is.na(d1$select)),]$select), main="Admit %")
plot(density(d1$stufacr), main="Student:faculty ratio")

table(d1$calsys,d1$control)

```

This code chunk recodes IPEDS' locale codes into something more readable.

```{r}
d1$locale2 <- substr(d1$locale,1,1)
d1$locale2 <- as.factor(d1$locale2)
levels(d1$locale2) <- c('City','Town','Suburb','Rural')
table(d1$locale2, d1$locale)
```


# Model and output

```{r}
theLM <- lm(rate ~ calsys + as.integer(tuition1) + control + select + stufacr + locale2  + adjusted_cohort, data=d1, family=gaussian)
summary(theLM)
library(gvlma)
gvlma(theLM)
plot(theLM)
vif(theLM)
```

